---
title: "Statistical Inference Course Project Part 1"
author: "Anas"
date: "2024-12-30"
output:
  pdf_document: default
  html_document: default
---
Synopsis
The project consists of two parts:
1. Simulation Exercise to explore inference
2. Basic inferential analysis using the ToothGrowth data in the R datasets package

**Part 1: Simulation Exercise**
The task is to investigate the exponential distribution (ED) in R and compare it with the Central Limit Theorem. The ED will be simulated in R with rexp(n,lambda) where lambda is the rate parameter. The mean of exponential distribution and the standard deviation are both 1/lambda where lambda = 0.2, and distribution of averages of 40 exponentials and will perform 1000 simulations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tinytex)

 set.seed(100)
lambda <- 0.2
n <- 40
sim_data <- replicate(1000, rexp(n, lambda))
mean_sim_data <- apply(sim_data, 2, mean)
```
**Mean Comparision**
*Sample Mean vs Theoretical Mean of the Distribution*
```{r}
# Sample Mean
sampleMean <- mean(mean_sim_data) # Mean of sample means
print (paste("Sample Mean = ", sampleMean))

# Theoretical Mean
# the expected mean of the exponential distribution of rate = 1/lambda
theoretical_mean <- (1/lambda)
print (paste("Theoretical Mean = ", theoretical_mean))
```
```{r}
# Histogram shows differences
hist(mean_sim_data, col="light blue", xlab = "Mean Average", main="Distribution of Exponential Average")
abline(v = theoretical_mean, col="brown")
abline(v = sampleMean, col="green")
```

*Sample variance vs Theoritical variance*
```{r}
# sample deviation & variance

sample_dev <- sd(mean_sim_data)
print (paste("Sample SD = ", sample_dev))

sample_variance <- sample_dev^2
print (paste("Sample V = ", sample_variance))
```

```{r}
# theoretical deviation & variance
theoretical_dev <- (1/lambda)/sqrt(n)
print (paste("Theoritical SD = ", theoretical_dev))
theoretical_variance <- ((1/lambda)*(1/sqrt(n)))^2
print (paste("Theoritical V = ", theoretical_variance))
```
*Distribution*
```{r}
d <- data.frame(mean_sim_data)
t <- data.frame(theoretical_mean)
g <- ggplot(d, aes(x = mean_sim_data)) +

 geom_histogram(binwidth = .2, color="black", fill="brown" , aes(y=..density..))+
            stat_function(fun=dnorm, args=list(mean=theoretical_mean, sd=sd(mean_sim_data)), 
                          color="green", size =1) +
            stat_density(geom = "line", color = "blue", size =1)  +
            labs(x="Mean", y= "Density", 
                    title="Normal Distribution Comparision")+
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12)
  )
g
```
The above plot indicated that density curve is somehow similar to normal distribution curve.
Q-Q Normal Plot also indicates the normal distribution
```{r}
qqnorm(mean_sim_data)
 qqline(mean_sim_data, col = "magenta")
```
